---
title: "HomeWork_5"
author: "Vincent Tamakloe"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1: Matrix Operations

### Assign a Random Integer for the Matrix
```{r}
n_dims <- sample(3:10, size=1)
n_dims
```

### Generate a Randomized Matrix
```{r}
# Create a sequence from 1 to n_dims^2
vector <- seq(1, n_dims^2)
# Shuffle the sequence to randomize the values
vector <- sample(vector)
# Convert the vector into a square matrix
matrix1 <- matrix(vector, nrow=n_dims)
print(matrix1)
```

### Apply a Matrix Transposition
```{r}
# Transpose the matrix so that rows become columns and vice versa
transposed_matrix <- t(matrix1)
print(transposed_matrix)
```

### Compute Basic Statistics on Specific Rows
```{r}
# Find the mean and sum for the first row
first_row_mean <- mean(matrix1[1,])
first_row_sum <- sum(matrix1[1,])
# Find the mean and sum for the last row
last_row_mean <- mean(matrix1[n_dims,])
last_row_sum <- sum(matrix1[n_dims,])

list(first_row_mean, first_row_sum, last_row_mean, last_row_sum)
```

### Perform Eigenvalue Analysis
```{r}
# Compute the eigenvalues and eigenvectors of the matrix
matrix_eigen <- eigen(matrix1)
print(matrix_eigen)
```

### Verify Data Types of Eigenvalues and Eigenvectors
```{r}
# Check the data types of eigenvalues and eigenvectors
typeof(matrix_eigen$values)
typeof(matrix_eigen$vectors)
```

## Task 2: Creating and Manipulating Lists

### Build a List Containing Different Data Types
```{r}
# Create a 4x4 matrix with random uniform values
my_matrix <- matrix(runif(16), nrow=4)
# Generate a logical vector of TRUE/FALSE values
my_logical <- (runif(100) > .8)
# Generate a randomized sequence of lowercase letters
my_letters <- sample(letters)
```

### Extract and Store Specific Elements in a New List
```{r}
# Create a new list extracting specific values from each component
list1 <- list(my_matrix, my_logical, my_letters)
new_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])
```

### Check Data Types of Extracted Elements
```{r}
# Verify the type of each extracted element
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
```

### Merge Elements into a Single Vector
```{r}
# Combine extracted elements into a single atomic vector
combined_elements_vector <- c(new_list[[1]], new_list[[2]], new_list[[3]])
print(combined_elements_vector)
# Determine the data type of the combined vector
typeof(combined_elements_vector)
```

## Task 3: Data Frames and Handling Missing Data

### Construct a Data Frame
```{r}
# Create a data frame with two columns
# First column: Random uniform values between 0 and 10
# Second column: Randomly ordered capital letters
dataframe <- data.frame(
  my_unis = runif(26, min=0, max=10), 
  my_letters = sample(LETTERS)
)
print(dataframe)
```

### Introduce Missing Values
```{r}
# Replace values in four randomly chosen rows with NA
dataframe[sample(1:26, size=4), 1] <- NA
```

### Identify Positions of Missing Data
```{r}
# Locate the rows where NA values appear
which(is.na(dataframe$my_unis))
```

### Reorder Data by Alphabetical Order
```{r}
# Arrange data based on the alphabetical order of the letter column
dataframe <- dataframe[order(dataframe$my_letters),]
print(dataframe)
```

### Compute the Mean of the Numeric Column
```{r}
# Calculate the mean of the numeric column, ignoring NA values
mean(dataframe$my_unis, na.rm=TRUE)
```
