---
title: "Home Work 8 - Simulating and Fitting Data Distributions"
author: "Vincent Tamakloe"
date: "2025-03-19"
output: "html_document"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading libraries
```{r}
library(ggplot2)
library(MASS)
library(readxl)
```

#### Reading My Data
```{r}
z <- read_excel("C:/Users/j/Desktop/FruitflyData.xlsx")  

# Rename the column
colnames(z)[colnames(z) == "Populations"] <- "myVar" 

# Check the structure and summary
str(z)
summary(z$myVar)
```

#### Plotting Histogram
```{r}
# Basic histogram of fruitfly population data
ggplot(z, aes(x = myVar)) +
  geom_histogram(color = "grey60", fill = "cornsilk", bins = 15) +
  labs(title = "Histogram of Fruitfly Counts",
       x = "Counts", y = "Frequency")
```

#### Plotting Histogram and Empirical Density
```{r}
p1 <- ggplot(z, aes(x = myVar, y = ..density..)) +
  geom_histogram(color = "grey60", fill = "cornsilk", size = 0.2, bins = 15) +
  geom_density(linetype = "dotted", size = 0.75)
p1
```

#### Fitting Normal Distribution
```{r}
normPars <- fitdistr(z$myVar, "normal")
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]
xval <- seq(0, max(z$myVar), length.out = length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour = "red",
                      args = list(mean = meanML, sd = sdML))
p1 + stat
```

#### Fitting Exponential Distribution
```{r}
expoPars <- fitdistr(z$myVar, "exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour = "blue",
                       args = list(rate = rateML))
p1 + stat + stat2
```

#### Fitting Uniform Distribution
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour = "darkgreen",
                       args = list(min = min(z$myVar), max = max(z$myVar)))
p1 + stat + stat2 + stat3
```

#### Fitting Gamma Distribution (Requires Positive Data)
```{r}
z_pos <- z[z$myVar > 0, ]
gammaPars <- fitdistr(z_pos$myVar, "gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour = "purple",
                       args = list(shape = shapeML, rate = rateML))
p1 + stat + stat2 + stat3 + stat4
```

#### Fitting Beta Distribution (On Rescaled Data)
```{r}
# Rescale data
z$scaledVar <- z$myVar / (max(z$myVar) + 0.1)

# Filter out values that are exactly 0 or 1
z_beta <- z[z$scaledVar > 0 & z$scaledVar < 1, ]

# Proceed with beta fit using cleaned data
pSpecial <- ggplot(z_beta, aes(x = scaledVar, y = ..density..)) +
  geom_histogram(color = "grey60", fill = "cornsilk", size = 0.2, bins = 15) +
  xlim(c(0, 1)) +
  geom_density(size = 0.75, linetype = "dotted")

betaPars <- fitdistr(z_beta$scaledVar, start = list(shape1 = 1, shape2 = 2), "beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

x_beta <- seq(0, 1, length.out = length(z_beta$scaledVar))

statSpecial <- stat_function(aes(x = x_beta, y = ..y..), fun = dbeta,
                             colour = "orchid",
                             args = list(shape1 = shape1ML, shape2 = shape2ML))
pSpecial + statSpecial
```

#### Simulating New Gamma Data
```{r}
set.seed(123)
sim_data <- rgamma(n = nrow(z_pos), shape = shapeML, rate = rateML)
sim_df <- data.frame(myVar = sim_data)

ggplot(sim_df, aes(x = myVar)) +
  geom_histogram(aes(y = ..density..), color = "grey60", fill = "cornsilk", bins = 15) +
  stat_function(fun = dgamma,
                args = list(shape = shapeML, rate = rateML),
                color = "brown") +
  labs(title = "Simulated Data from Gamma Distribution",
       x = "Simulated Counts", y = "Density")
```

#### Comparing Histogram of Original Data with Gamma Fit
```{r}
ggplot(z_pos, aes(x = myVar)) +
  geom_histogram(aes(y = ..density..), color = "grey60", fill = "cornsilk", bins = 15) +
  stat_function(fun = dgamma,
                args = list(shape = shapeML, rate = rateML),
                color = "brown") +
  labs(title = "Original Data with Gamma Fit",
       x = "Observed Counts", y = "Density")
```

#### Q3. How do the two histogram profiles compare?
The simulated data and the original fruitfly population counts both show a strongly right-skewed distribution. Their histograms are similar in shape and spread, with the gamma curve aligning well to both datasets. The simulated data captures the main features — central tendency, spread, and skew — of the real data.

#### Q4. Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?
Yes, the gamma model does a good job. This is because:

i. The fitted curve overlays well on the histogram of the original data.

ii. The simulated data have a similar shape and range as the original.

iii. Gamma distributions are flexible for modeling positive, right-skewed count data like yours.

Minor differences in the tails might occur, but overall, the gamma model is both statistically and visually a strong fit.
