---
title: "Home Work 5 - Matrices, Lists, and Data Frames"
author: "Vincent Tamakloe"
date: "2025-02-12"
output: "html_document"
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Task 1: Matrix Operations 

##### a. Generate matrix dimensions
```{r}
n_dims <- sample(3:10, size=1)
n_dims
```

##### b. Create randomized square matrix
```{r}
vector <- sample(seq(1, n_dims^2))
matrix1 <- matrix(vector, nrow = n_dims)
print(matrix1)
```

##### c. Transpose matrix
```{r}
transposed_matrix <- t(matrix1)
print(transposed_matrix)
```

##### d. Compute Basic Statistics on Specific Rows
```{r}
first_row_mean <- mean(matrix1[1, ])
first_row_sum <- sum(matrix1[1, ])
last_row_mean <- mean(matrix1[n_dims, ])
last_row_sum <- sum(matrix1[n_dims, ])
list(first_row_mean, first_row_sum, last_row_mean, last_row_sum)
```

##### e. Perform Eigenvalue Analysis
```{r}
matrix_eigen <- eigen(matrix1)
print(matrix_eigen)
```

##### f. Check Types of Eigenvalues and Eigenvectors
```{r}
typeof(matrix_eigen$values)
typeof(matrix_eigen$vectors)
```


#### Task 2: Creating and Manipulating Lists

##### a. Build a List Containing Different Data Types
```{r}
# Create a 4x4 matrix with random uniform values
my_matrix <- matrix(runif(16), nrow=4)

# Generate a logical vector of TRUE/FALSE values
my_logical <- (runif(100) > 0.8)

# Generate a randomized sequence of lowercase letters
my_letters <- sample(letters)
```

##### b. Extract and Store Specific Elements into a New List
```{r}
# Combine the objects into one list
list1 <- list(my_matrix, my_logical, my_letters)

# Extract the second element from each and store in a new list
new_list <- list(my_matrix[2, 2], my_logical[2], my_letters[2])
```

##### c. Check Data Types of Extracted Elements
```{r}
# Inspect the type of each item in new_list
typeof(new_list[[1]])  # numeric
typeof(new_list[[2]])  # logical
typeof(new_list[[3]])  # character
```

##### d. Merge Elements into a Single Vector
```{r}
# Combine all extracted values into one atomic vector
combined_elements_vector <- c(new_list[[1]], new_list[[2]], new_list[[3]])

# View the merged vector and its type
print(combined_elements_vector)
typeof(combined_elements_vector)
```


#### Task 3: Data Frames and Handling Missing Data

##### a. Construct Data Frame
```{r}
# Create a data frame with two columns:
# - Random uniform values between 0 and 10
# - Randomly ordered capital letters
dataframe <- data.frame(
  my_unis = runif(26, min = 0, max = 10),
  my_letters = sample(LETTERS)
)

print(dataframe)
```

##### b. Introduce Missing Values
```{r}
# Replace four values in the numeric column with NA
dataframe[sample(1:26, size = 4), 1] <- NA
```

##### b. Identify Positions of Missing Data
```{r}
# Locate the rows where NA values appear
which(is.na(dataframe$my_unis))
```

### Reorder Data by Alphabetical Order
```{r}
# Sort the data frame alphabetically by the letter column
dataframe <- dataframe[order(dataframe$my_letters), ]
print(dataframe)
```

### Compute the Mean of the Numeric Column
```{r}
# Calculate the mean of the numeric column, excluding NAs
mean(dataframe$my_unis, na.rm = TRUE)
```
