---
title: "HomeWork_8"
author: "Vincent Tamakloe"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#### Load required libraries
```{r}
library(tidyverse)
```

#### Load the iris dataset
```{r}
data(iris) 
```



#### Question 1. Examine the structure of the iris dataset
```{r}
glimpse(iris) 
```



#### Question 2. Create iris1: Filter for species Virginica and Versicolor with Sepal.Length > 6 and Sepal.Width > 2.5
```{r}
iris1 <- iris %>%
  filter(Species %in% c("virginica", "versicolor"),
         Sepal.Length > 6,
         Sepal.Width > 2.5)

glimpse(iris1)
```



#### Question 3. Create iris2: Select only Species, Sepal.Length, and Sepal.Width
```{r}
iris2 <- iris1 %>%
  select(Species, Sepal.Length, Sepal.Width)

glimpse(iris2)
```



#### Question 4. Create iris3: Arrange by Sepal.Length in descending order
```{r}
iris3 <- iris2 %>%
  arrange(desc(Sepal.Length))

head(iris3)
```



#### Question 5. Create iris4: Add a column for Sepal.Area (Sepal.Length * Sepal.Width)
```{r}
iris4 <- iris3 %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width)

glimpse(iris4)
```



#### Question 6. Create iris5: Compute overall mean Sepal.Length, mean Sepal.Width, and sample size
```{r}
iris5 <- iris4 %>%
  summarise(Avg.Sepal.Length = mean(Sepal.Length),
            Avg.Sepal.Width = mean(Sepal.Width),
            Sample.Size = n())

print(iris5)
```



#### Question 7. Create iris6: Compute mean Sepal.Length, mean Sepal.Width, and sample size grouped by Species
```{r}
iris6 <- iris4 %>%
  group_by(Species) %>%
  summarise(Avg.Sepal.Length = mean(Sepal.Length),
            Avg.Sepal.Width = mean(Sepal.Width),
            Sample.Size = n())

print(iris6)
```



#### Question 8. Use piping to create irisFinal in one pipeline
```{r}
irisFinal <- iris %>%
  filter(Species %in% c("virginica", "versicolor"),
         Sepal.Length > 6,
         Sepal.Width > 2.5) %>%
  select(Species, Sepal.Length, Sepal.Width) %>%
  arrange(desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width)

head(irisFinal)
```



#### Question 9. Create a longer data frame with columns: Species, Measure, Value
```{r}
iris_long <- iris %>%
  pivot_longer(cols = -Species, names_to = "Measure", values_to = "Value")

print(iris_long)
```

